---
title: "Lab 9: Baby Names"
author: "Hannah Pawig"
format:
  html:
    theme: lux
    smooth-scroll: true
    code-fold: true
    code-summary: 'This is where the code lives'
editor: visual
execute:
  error: false
---

```{r}
#| include: false
#| message: false
library(tidyverse)
library(here)
library(ggtext)
options(dplyr.summarise.inform = FALSE)

```

```{r data}
#| message: false
A_baby_names <- read_csv(here('Week 9',
                            'data',
                            'StateNames_A.csv'))
```

# Summarizing & Visualizing the Number of Allisons

## 1. Summary Table

> Number of babies named "Allison" for each state and sex of the baby.

```{r}
A_baby_names |>
  rename(Sex = 'Gender') |> 
  filter(Name == 'Allison') |> 
  group_by(State, Sex) |> 
  summarise(
    Num_Allisons = sum(Count)
  )  |>
  pivot_wider(
              names_from = Sex,
              values_from = Num_Allisons) |> 
  mutate(
    across(.cols = 'F':'M',
           .fns = ~replace_na(., 0))
  )
```

## 2. Focusing on babies assigned as Female at birth

```{r}
#| code-fold: false
Allisons <- A_baby_names |> 
  rename(Sex = 'Gender') |> 
  filter(Name == 'Allison',
         Sex == 'F') |> 
  group_by(Year) |> 
  summarise(Num_Allisons = sum(Count))

```

## 3. Visualization

```{r}
Allisons |> 
  ggplot(mapping = aes(x = Year,
                       y = Num_Allisons)) +
  geom_line(color = 'deeppink') +
  labs(title = 'Number of Female Babies named Allison',
       subtitle = 'United States, 1997-2014') +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7)) +
  scale_x_continuous(
    breaks = seq(from = 1997, to = 2014, by = 1)
  )

```

# Modeling the Number of Allisons

## 4. Creating a linear model with the year as the explanatory variable and the number of Allisons as the response variable.

```{r}
#| message: false
#| code-fold: false
lm_Allisons <- Allisons |> 
  lm(formula = Num_Allisons ~ Year,
     data = _)
```

## 5. Visualize the Regression Model

```{r}
#| message: false
lm_Allisons |> 
  ggplot(mapping = aes(x = Year,
                       y = Num_Allisons)) +
  geom_point() +
  stat_smooth(method = 'lm') +
  labs(title = 'Time (Years) vs. Number of Babies named "Allison"',
       x = '\nTime (Years)') +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 10),
    axis.text.x = element_text(size = 7)
  ) +
  scale_x_continuous(
    breaks = seq(from = 1997, to = 2014, by = 1)
  )
```

## 6. The Estimated Regression Equation

```{r}
lm_Allisons
```

Predicted number of babies named Allison = 209689.8 - 101.5(Year)

## 7. Plotting Residuals

```{r}
lm_Allisons |> 
  broom::augment() |> 
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point() +
  labs(title = 'Predicted vs. Residual Number of Allisons',
       subtitle = 'Estimated based on a given year, 1997-2014') +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

> No, I don't see any obvious patterns in the residual plot, the points are just randomly scattered around 0.

## 8. What can we conclude from the model?

Per year increase, there are 101 fewer Allisons than the previous year, and overall there is a downward trend for naming babies "Allison" from 1997 to 2014.

# Spelling by State

## 9. Finding the Allan/Alan/Allen babies

```{r}
Als <- A_baby_names |> 
  rename(Sex = 'Gender') |> 
  filter(Name %in% c('Allan', 'Allen', 'Alan'),
         Sex == 'M')


Als |> 
  group_by(Year, Name) |> 
  summarise(
    Num_Als = sum(Count)
  ) |> 
  ggplot(mapping = aes(x = Year,
                       y = Num_Als,
                       color = Name)) +
  geom_line() +
  labs(title = "Babies named <span style='color:red'>Alan</span>, <span style='color
                :darkorange'>Allan</span>, or <span style='color:chocolate4'>Allen</span>",
       subtitle = 'United States, 1997-2014') +
  scale_color_manual(values = c('red', 'darkorange', 'chocolate4')) +
  theme(legend.position = 'none',
        plot.title = element_markdown(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7)) +
  scale_x_continuous(breaks = seq(from = 1997, to = 2014, by = 1))

```

## 10. The Allan babies in California and Pennsylvania (in 2000)

```{r}

Als |> 
  filter(State %in% c('CA', 'PA'),
         Year == 2000) |>
  group_by(State, Name) |> 
  summarise(
    Count = sum(Count)
  ) |> 
  mutate(
     Count = replace_na(Count, 0)
  )|> 

  pivot_wider(names_from = Name,
            values_from = Count)

```

## 11. Percentage Breakdown for the 3 spellings of *Allan* in CA & PA (2000)

```{r calculate-total-function}
#| code-summary: 'Creating a calculate_total function'
calculate_total <- function(df, state, year) {
 df <- df |> 
    filter(Year == {{ year }},
           State == {{ state }})
 total <- sum(df$Count)
   
  return(total)
}
```

```{r}
Als |> 
  filter(State %in% c('CA', 'PA'),
         Year == 2000) |>
  group_by(State, Name) |>
    mutate(
     Count = replace_na(Count, 0),
     Total = case_when(
        State == 'CA' ~ calculate_total(df = Als, state = 'CA', year = 2000),
        State == 'PA' ~ calculate_total(df = Als, state = 'PA', year = 2000))
      ) |> 
  summarise(
    Percentage = round((sum(Count) / mean(Total)) * 100,
                        digits = 2)) |>
  pivot_wider(names_from = Name,
            values_from = Percentage)
```
