---
title: "Lab 4"
author: "Hannah Pawig"
format: 
  html:
    smooth-scroll: true
    code-fold: true
    code-summary: "Click me to see the code!"
    theme: lux 
    self-contained: true
execute:
  message: false
editor: visual
---

```{r setup}
#| message: false
library(tidyverse)
library(here)
library(scales)
avocados <- read_csv(here::here("supporting_artifacts",
                                "learning_targets",
                                "Lab 4 NonRevised",
                                "avocado.csv"))
```

# Dataset Description

The data set contains information about the weekly total retail volume (units) of Hass avocados throughout the nation, from 2015 to 2018. Additionally, the data set holds average price for an avocado, whether the avocado is organic or not, and the total number of Hass avocados sold for each PLU code: PLU 4046, PLU 4225, and PLU 4770, which correspond to the Hass avocado size (small, medium/large, and extra large respectively). This data set also includes information about the number of bags the avocados were sold in, and the sizes of the bags they were sold in.

```{r data_cleaning}
avocados_clean <- avocados |>
  #fixing column names
  rename(total_volume = `Total Volume`,
         average_price = `AveragePrice`,
         total_bags = `Total Bags`,
         date = `Date`)

# creating helper data sets
avocados_region <- data.frame(region = c("Midsouth", "GreatLakes",
                                         "Northeast",
                                         "SouthCentral", "Southeast",
                                         "West", "WestTexNewMexico"))

avocados_state <- data.frame(state = "California")

avocados_totalUS <- data.frame(region = "TotalUS")
```

# Exercises {style="color: yellowgreen"}

## 3) Which major region sold the most organic, small Hass avocados in 2017? {style="color: darkolivegreen"}

```{r exercise_pt_1}
avocados_clean |> 
  pivot_longer(cols = `4046`:`4770`,
               names_to = "PLU",
               values_to = "vol_by_PLU") |>
  select(region, date, total_volume, year, type, PLU, vol_by_PLU) |>
  semi_join(avocados_region, by = "region") |>
  filter(year == "2017",
         type == "organic",
         PLU == "4046") |>
  slice_max(order_by = vol_by_PLU)

```

The Northeast region sold the most organic, small Hass avocados in 2017, in the week of February 19th. The region's volume sold of this specific avocado type was 127,051.75 units.

## 4) In which month is the highest volume of avocado sales? {style="color: saddlebrown"}

```{r exercise_pt_2}
# resource used: https://www.statology.org/separate-function-in-r/

# separating date observations into year, month, and day
avocados_clean |>
  semi_join(avocados_region, by = "region") |>
  separate(col = date,
           into = c('year','month','day'),
           sep = '-') |>
  group_by(month) |>
  summarise(sum_total_volume = sum(total_volume)) |>
  slice_max(order_by = sum_total_volume)
```

The highest volume of avocados sold in a month was 533,030,497 avocados, which occurred in January.

## 5) Which metro area regions sold the most total avocados? {style="color: darkolivegreen"}

Metro area regions that sold the *most* total avocados, based on their respective average total volume across the years 2015 to 2018 were

```{r finding_metro_areas}
# using helper data sets to isolate the metro areas
avocados_metro <- avocados_clean |>
  anti_join(avocados_state, by = c('region' = 'state')) |>
  anti_join(avocados_region, by = c('region' = 'region')) |>
  anti_join(avocados_totalUS, by = c('region' = 'region')) |>
  rename(metro_area = `region`)

# finding the top 5 metro areas with highest avg total volume
metro_top_5 <- avocados_metro |>
  select(metro_area, total_volume, year) |>
  group_by(metro_area) |> # group by city in order to calculate total volume means                           # within each city
  summarise(mean_total_vol = mean(total_volume)) |>
  arrange(desc(mean_total_vol)) |>
  slice_head(n = 5)
```

1.  Los Angeles (California)
2.  Plains (Georgia)
3.  New York (New York)
4.  Dallas-Fort Worth (Texas)
5.  Houston (Texas)

## Let's visualize! {style="color: saddlebrown"}

```{r boxplots}
# resources used for changing axis intervals:
# https://www.geeksforgeeks.org/how-to-change-axis-intervals-in-r-plots/
# https://www.datanovia.com/en/blog/ggplot-axis-ticks-set-and-rotate-text-labels/

#| fig.align: 'center'
#| fig.cap: 'Only total volumes of the top 5 metro areas with the highest overall average total volume are shown'

avocados_metro |>
  semi_join(metro_top_5, by = c("metro_area" = "metro_area")) |>
  mutate(total_volume = total_volume / 1000) |> #changing units of total volume to                                                            thousand units
  ggplot(mapping = aes(x = total_volume,
                       y = metro_area,
                       alpha = 0.0003)) +
  geom_boxplot() +
  geom_jitter(color = "seagreen") +
  scale_x_continuous(breaks = seq(from = 0, to = 5500, by = 250)) +
  theme(axis.text.x = element_text(size = 5)) +
  labs(title = 'Metro area regions & Total Volume of Sold Avocados',
       x = 'Total Volume (by 1,000 units)',
       y = 'Metro Area')

```

# Reshaping {style="color: yellowgreen"}

> We're going to answer some questions about regions in California.

## 6) In which regions is the price of organic versus conventional avocados most different? {style="color: darkolivegreen"}

```{r}
#| message: false

#helper data frame for only CA cities
CA_cities <- data.frame(city = c('LosAngeles', 'SanDiego',
                                  'Sacramento', 'SanFrancisco'))


avocados_california_regions <- avocados_metro |>
  semi_join(CA_cities, by = c("metro_area" = "city")) |>
  group_by(metro_area, type) |>
  mutate(avg_price_by_type = mean(average_price)) |>
  distinct(metro_area, .keep_all = TRUE)

avocados_california_regions |>
  select(metro_area, type, avg_price_by_type) |>
  group_by(metro_area) |>
  summarise(
    diff_price = round(abs(
                            avg_price_by_type[match("conventional", type)] -
                            avg_price_by_type[match("organic", type)]), digits = 2))

# tibble of avg prices by CA region and avocado type
avocados_california_regions |>
  group_by(metro_area) |>
  select(metro_area, type, avg_price_by_type)



```

Los Angeles difference in means of conventional vs. organic ≈ \$0.48

Sacramento difference in means of conventional vs. organic ≈ \$0.70

San Diego difference in means of conventional vs. organic ≈\$0.67

San Francisco difference in means of conventional vs. organic ≈ \$0.81

```{r plot_price_diffs}
avocados_california_regions |>
  ggplot(mapping = aes(x = type, y = avg_price_by_type, fill = type)) +
  geom_col() +
  facet_wrap(~metro_area) +
  labs(fill = "Avocado Type",
       x = "Type",
       y = "Average Price (USD)")
```

***San Francisco is the region where the difference in price of organic vs. conventional is largest.***

> Visually, the difference in height (height representing the average price) between the conventional and organic columns is largest in the San Francisco graph, with the difference looking to be about 0.65.

## 7) Plot of California Regions {style="color: saddlebrown"}

```{r making_data_set_w_props}
avocados_california_regions_props <- avocados_california_regions |>
  select(metro_area, type, total_volume, `4046`, `4225`, `4770`) |>
  pivot_longer(cols = `4046`:`4770`,
               names_to = "PLU",
               values_to = "volume_by_PLU") |>
  group_by(metro_area, PLU, type) |>
  mutate(
    proportion = volume_by_PLU / total_volume
  )|>
  distinct(metro_area, .keep_all = TRUE)
```

```{r prop_plot}
# resource for x axis tick labels
# https://stackoverflow.com/questions/13297995/changing-font-size-and-direction-of-axes-text-in-ggplot2

#legend labels
# https://www.statology.org/ggplot-legend-labels/
# https://www.statology.org/change-legend-title-ggplot2/

avocados_california_regions_props |>
  ggplot(mapping = aes(x = metro_area, 
                       y = proportion, 
                       fill = PLU)) +
  geom_col(position = "fill") +
  facet_wrap(~type) +
  labs(x = "Region of CA",
       y = "Proportion of Mean Avocados Sold",
       fill = "Avocado Size") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_fill_discrete(labels = c('Small', 'Large', 'Extra Large'))
```
